# Reproducible Research: Peer Assessment 1



## Loading and preprocessing the data

First, we load the data:

```{r, echo=TRUE, results='markup'}
data <- read.table("activity.csv", header=TRUE, sep=",")
str(data)
```

Then, we convert the date variable to date type:

```{r}
data <- transform(data, date = as.Date(date))
str(data)
head(data)
```


## What is mean total number of steps taken per day?

First, we remove the missing values:

```{r}
completeCases <- data[complete.cases(data$steps),c("steps","date")]
```

Then, we calculate the number of steps taken per day and rename the columns:

```{r}
stepsPerDay <- aggregate(completeCases$steps, by=list(completeCases$date), FUN=sum)
names(stepsPerDay) <- c("date", "steps")
```

Then, we make a histogram of the total number of steps taken per day:

```{r}

barplot(
    stepsPerDay$steps, 
    names.arg=stepsPerDay$date,
    main="Total steps taken per day",
    xlab="Day",
    ylab="Steps",
    col="red")
```

An finally, we calculate the mean and the median total number of steps taken per day:

```{r}
meanSteps <- mean(stepsPerDay$steps)
medianSteps <- median(stepsPerDay$steps)
```

- The mean number of steps taken per day is **`r meanSteps`**.
- The median number of steps taken per day is **`r medianSteps`**.


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
