# Reproducible Research: Peer Assessment 1



## Loading and preprocessing the data

First, we load the data:

```{r, echo=TRUE, results='markup'}
data <- read.table("activity.csv", header=TRUE, sep=",")
str(data)
```

Then, we convert the date variable to date type:

```{r}
data <- transform(data, date = as.Date(date))
str(data)
head(data)
```


## What is mean total number of steps taken per day?

First, we remove the missing values:

```{r}
completeCases <- data[complete.cases(data$steps),c("steps","date")]
```

Then, we calculate the number of steps taken per day and rename the columns:

```{r}
stepsPerDay <- aggregate(completeCases$steps, by=list(completeCases$date), FUN=sum)
names(stepsPerDay) <- c("date", "steps")
```

Then, we make a histogram of the total number of steps taken per day:

```{r}

barplot(
    stepsPerDay$steps, 
    names.arg=stepsPerDay$date,
    main="Total steps taken per day",
    xlab="Day",
    ylab="Steps",
    col="red")
```

An finally, we calculate the mean and the median total number of steps taken per day:

```{r}
meanSteps <- mean(stepsPerDay$steps)
medianSteps <- median(stepsPerDay$steps)
```

- The mean number of steps taken per day is **`r meanSteps`**.
- The median number of steps taken per day is **`r medianSteps`**.


## What is the average daily activity pattern?

First, we remove the missing values:

```{r}
completeCases <- data[complete.cases(data$steps),]
```

Then, we calculate the average number of steps taken per interval and rename the columns:

```{r}
stepsPerInterval <- aggregate(
        completeCases[,c("steps")], 
        by=list(completeCases$interval), 
        FUN=mean)

names(stepsPerInterval) <- c("interval", "steps")

```

Then, we make a time series plot with the average daily activity: 

```{r}
plot(
    stepsPerInterval$interval, 
    stepsPerInterval$steps, 
    type="l",
    main="Average daily activity",
    xlab="5-minute interval",
    ylab="Average number of steps")
```

Finally, we calculate the 5-minute interval wich contains the maximum number of steps.

```{r}
IntervalMaxSteps <- stepsPerInterval[order(-stepsPerInterval$steps),][1,"interval"]
```

The interval **`r IntervalMaxSteps`** is the 5-minute interval wich contains the maximum number of steps.


## Imputing missing values

First, we calculate the total number of missing values in the dataset:

```{r}
totalMissing <- nrow(data[is.na(data),])
```

The total number of missing values in the dataset is **`r totalMissing`**.

Then, we are going to create a dataset with the missing values filled with the mean for that 5-minute interval:

```{r}
dataFilled <- data

for (i in 1:nrow(dataFilled))
{
    if(is.na(dataFilled$steps[i]))
    {
        interval <- stepsPerInterval$interval == dataFilled$interval[i]
        dataFilled$steps[i] = stepsPerInterval[interval, "steps"]
    }
}
```

Then, we calculate the number of steps taken per day and rename the columns:

```{r}
stepsPerDayFilled <- aggregate(
    dataFilled$steps, 
    by=list(dataFilled$date), 
    FUN=sum)

names(stepsPerDayFilled) <- c("date", "steps")
```

Then, we make a histogram of the total number of steps taken per day:

```{r}

barplot(
    stepsPerDayFilled$steps, 
    names.arg=stepsPerDayFilled$date,
    main="Total steps taken per day",
    xlab="Day",
    ylab="Steps",
    col="red")
```

An finally, we calculate the mean and the median total number of steps taken per day:

```{r}
meanStepsFilled <- mean(stepsPerDayFilled$steps)
medianStepsFilled <- median(stepsPerDayFilled$steps)
```

- The mean number of steps taken per day is **`r meanStepsFilled`**.
- The median number of steps taken per day is **`r medianStepsFilled`**.

If we compare this values with the values calculated in the first part, we can see that **the mean has the same value but the median has slightly increased**.


## Are there differences in activity patterns between weekdays and weekends?




